---
title: "Careless Responding"
subtitle: "Title of your Analysis"
author: 
 - name: Björn S. Siepe
   orcid: 0000-0002-9558-4648
   affiliations: University of Marburg
 - name: Add name(s) of person(s) responsible for respective file
   orcid: 0000-0002-9558-4648
   affiliations: to be added
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    code-summary: "Show the code"
    fig-width: 7
    fig-height: 4.5
    fig-align: "center"
    embed-resources: true
execute:
  message: false
  warning: false
  eval: true # run code?
---

# Background
You can explain your analyses here. 


We first load all relevant packages: 
```{r packages}
if(!require(pacman)) install.packages("pacman")
pacman::p_load("dplyr", "tidyr", "here", "ggplot2")  
  
set.seed(35032)  
```


Some good practice: 

-  when loading data, use the `here::here()` function to avoid (a) absolute paths and (b) issues with different usage of '/' and '\\' across systems:

```{r, eval = FALSE}
data <- readRDS(here::here("data", "cleaned_data", "data.RDS"))
```



# Data Preparation

This section should contain all relevant preprocessing you need to do for your analyses. 
```{r data-preparation}
# code for data cleaning etc.
```



# Analysis

You can create multiple subsections for your analyses. 

## Analysis 1

## Analysis 2

Just including a mock plot here to show how it would look like within a .qmd: 
```{r data-analysis}
# code for analyses

# create a mock plot to illustrate how this would look in quarto:
data <- rnorm(10000)

data |> 
  as.data.frame() |> 
  ggplot(aes(x = data)) + 
  geom_density() + 
  theme_classic()
```




# Obtain standard R script

We can also obtain a standard `.R` script from this document by running the following code (assuming there is a folder called `\scripts\`): 

```{r purl-script, eval = FALSE}
knitr::purl(here::here("scripts", "title_of_your_script.qmd"), 
            output = here::here("scripts", "title_of_your_script.R"),
            documentation = 2)
```

