---
title: "Careless Responding"
subtitle: "Title of your Analysis"
author: 
 - name: Björn S. Siepe
   orcid: 0000-0002-9558-4648
   affiliations: University of Marburg
 - name: Add name(s) of person(s) responsible for respective file
   orcid: 0000-0002-9558-4648
   affiliations: to be added
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    code-summary: "Show the code"
    fig-width: 7
    fig-height: 4.5
    fig-align: "center"
    embed-resources: true
execute:
  message: false
  warning: false
  eval: true # run code?
---

# Background
In this file, we will compute all indices of careless responding. We aim for the following structure: 
`external_id`, `counter`, `index`. There should be no implicit missingness. 


We first load all relevant packages: 
```{r packages}
if(!require(pacman)) install.packages("pacman")
pacman::p_load("dplyr", "tidyr", "here", "ggplot2", "rrcov", "Hmisc")  

source(here("scripts", "00_functions.R"))
  
set.seed(35032)  
```
We load the data and remove all non-EMA prompts: 
```{r read-data}
data <- read.csv(here::here("data", "WARND_stage2_c1234_trainingset_before_30min_v2024_07_16.csv"))

data <- data |> 
  filter(!is.na(counter))
```

Relevant EMA items and their reaction times: 
```{r ema-itemss}
ema_items <-  c("sad_d", "stressed_d", "overwhelm_d", "nervous_d", "ruminate_d", "irritable_d", "tired_d", "cheerful_d", "motivated_d", "relaxed_d")
rt_ema_items <- paste0("rt_", ema_items)

```


# Data Preparation

We first subset the data for testing the functions on a smaller dataframe: 

```{r subset-data}
# select 10 individuals at random
ids_to_subset <- data |> 
  distinct(external_id) |> 
  slice_sample(n = 10) |> 
  pull(external_id)


subset_data <- data |> 
  filter(external_id %in% ids_to_subset)
```



# Analysis

Currently, this still depends on `sd_within_assessment` being the first, to have no implicit missingness when using `left_join()` later below.
```{r}
# Define the functions and their parameters
index_functions <- list(
    sd_within_assessment = list(
    func = calc_within_assessment_sd,
    args = list(
      items = ema_items,
      id_col = "external_id"
    )
  ),
  mahalanobis = list(
    func = calc_indicator_mahalanobis,
    args = list(items = ema_items)
  ),
  robust_pca = list(
    func = calc_indicator_rob_PCA_orthogonal_distance,
    args = list(items = ema_items)
  ),
  psychometric_syntonym = list(
    func = calc_psychometric_synonym_violations,
    args = list(
      items = ema_items,
      cor_threshold = 0.45,
      synonym_difference_threshold = 5
    )
  ),
  psychometric_antonym = list(
    func = calc_psychometric_antonym_violations,
    args = list(
      items = ema_items,
      cor_threshold = 0.45,
      antonym_maxvalue_threshold = 5
    )
  ),
  sd_response_times = list(
    func = calc_sd_response_times,
    args = list(items = rt_ema_items)
  ),

  items_at_mode = list(
    func = calc_mode_percentage,
    args = list(
      items = ema_items,
      id_col = "external_id"
    )
  ),
  longstring = list(
    func = calc_longstring_proportions, 
    args = list(
      items = ema_items, 
      rt_items = rt_ema_items
    )
  )
)


```


Calculate them all: 

```{r calc-indices}
# calculate all indices
indices <- index_functions |>
  purrr::map(
    .f = function(x) {
      do.call(x$func, c(list(data = subset_data), x$args))
    }
  )

# combine indices
df_indices <- indices |>
  purrr::reduce(
    .f = function(x, y) {
      x |>
        left_join(y, by = c("external_id", "counter"))
    }
  ) |> 
  dplyr::mutate(mode_pct = if_else(is.nan(mode_pct), NA, mode_pct))
```



# Visualizations







# Session Info

```{r session-info}
pander::pander(sessionInfo())

```

