---
title: "Online Supplement 1: Visualizing Indices"
subtitle: "For the Manuscript: Identifying Careless Responding in Ecological Momentary Assessment: Inconsistent Signals from Different Detection Methods in the WARN-D Data"
author: 
 - name: Björn S. Siepe
   orcid: 0000-0002-9558-4648
   affiliations: University of Marburg
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    code-summary: "Show the code"
    fig-width: 8
    fig-height: 5
    fig-align: "center"
    embed-resources: true
execute:
  message: false
  warning: false
  eval: true # run code?
params:
  rerun: false  # define parameter if all large analyses should be rerun
---

# Background

In this document, we visualize and analyze all indices that were computed in `01_calculate_indices.qmd` *without* the reaction time (RT) indices.

We first load all relevant packages: 
```{r packages}
if(!require(pacman))
  install.packages("pacman")
pacman::p_load(
  "here",
  "ggplot2",
  "rrcov",
  "Hmisc",
  "sysfonts",
  "showtext",
  "ggh4x",
  "cowplot",
  "gtsummary",
  "gt",
  "future",
  "furrr",
  "dplyr",
  "tidyr",
  "vegan",
  "tibble",
  "purrr",
  "flextable",
  "patchwork"
)

source(here("scripts", "00_functions.R"))

# use color based on the Johnson palette from MetBrewer
fill_color <- "#132b69"

set.seed(35032)  
```


Relevant EMA items and their reaction times: 
```{r ema-itemss}
ema_items <-  c("sad_d", "stressed_d", "overwhelm_d", "nervous_d", "ruminate_d", "irritable_d", "cheerful_d", "motivated_d", "relaxed_d")
rt_ema_items <- paste0("rt_", ema_items)
```


For later visualizations, we also load the results of different model-based approaches. Here, we use the LPA without RT indices and the mixture IRT model:
```{r load-models}
lpa_res_no_rt <- readRDS(here::here("output", "data_profiles_without_time_indices.RDS"))
irt_res <- readRDS(here::here("output", "data_mixtureIRT.RDS"))
```

We load other results of the indices:
```{r}
#| label: load-flags
flag_df_no_rt <- readRDS(here::here("output", "flag_df_no_rt.RDS"))

```




### Multiverse Similarity
Instead of checking the correlation between affect items, we can instead check the overlap between carelessness flags based on the indices and the model-based indices across the multiverse. To do so, we use similarity metrics.

Visualize the results for both models and the single and double hurdle approach: 
```{r}
#| label: multiverse-similarity-plots
df_jaccard_spec_no_rt <- readRDS(here::here("output", "df_jaccard_spec_no_rt.RDS"))
hurdles <- c("single", "double")
models <- c("lpa", "irt")
type <- "similarity"

# plots for similarities
for(model in models) {
  for(hurdle in hurdles) {
    sim_plot <- plot_similarity(df_jaccard_spec_no_rt , model, hurdle = hurdle)
    spec_plot <- plot_specification(df_jaccard_spec_no_rt , type = "similarity", model = model, hurdle = hurdle)
    
    mv_plot <- plot_grid(
      sim_plot,
      spec_plot,
      ncol = 1,
      rel_heights = c(1, 2),
      align = "h",
      axis = "b"
    )
    
    ggsave(
      paste0("mv_plot_similarity_no_rt", model, "_", hurdle, ".pdf"),
      mv_plot,
      device = "pdf",
      path = here::here("figures"),
      height = 10,
      width = 10
    )
  }
}
```



We can also check the highest similarity indices to the LPA across all specifications:
```{r}
#| label: multiverse-similarity-highest

df_jaccard_spec_no_rt |> 
  arrange(desc(jaccard_similarity1_lpa)) |> 
  head(10)
```



# Similarity Indices and Models

We attach the model results to the indices. We classify everyone with an LPA profile 1, and a mixture IRT class 1 as attentive. 
```{r}
#| label: attach-model
df_indices_models <- lpa_res_no_rt |> 
  left_join(irt_res, by = c("external_id", "counter")) |> 
  left_join(flag_df_no_rt, by = c("external_id", "counter")) |> 
    mutate(across(where(is.logical), ~ as.integer(.))) |> 
    mutate(flag_lpa = ifelse(LPA_profile != 1, 1, 0),
           flag_irt = ifelse(mixtureIRT == 1, 0, 1)) |> 
   select(contains("flag")) |> 
  # recode to obtain single hurdle and multiple hurdle approach
  mutate(
    single_hurdle = if_else(flag_sum == 0, 0, 1),
    double_hurdle = if_else(flag_sum < 2, 0, 1)
  ) |> 
  # then remove raw sum
  select(!flag_sum)
  
```




# Session Info

```{r session-info}
pander::pander(sessionInfo())

```

